---
title: "Visualization"
output:
  github_document:
    toc: true
author: Guojing Wu
data: 2018-09-27
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( # default setting
  echo = TRUE, 
  fig.width = 6, 
  fig.asp = 0.6, 
  out.width = "90%")
library(tidyverse)
library(rnoaa)
library(ggridges)
```

data import

```{r data import, cache=TRUE}
if (!weather_df) {
  weather_df = 
    rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                        var = c("PRCP", "TMIN", "TMAX"), 
                        date_min = "2017-01-01",
                        date_max = "2017-12-31") %>%
    mutate(
      name = recode(id, USW00094728 = "CentralPark_NY", 
                        USC00519397 = "Waikiki_HA",
                        USS0023B17S = "Waterhole_WA"),
      tmin = tmin / 10,
      tmax = tmax / 10) %>%
    select(name, id, everything())
}
```

## start a plot

blank

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

filter

```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

or just save it into a variable

```{r}
weather_sp = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

### more options

color

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name))
```

add a geom

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE)
```

make it transparent

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = 0.4)) +
  geom_smooth(se = FALSE)
```

so what will happen if I put color parameters in ggplot()? It became a global parameter

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

now we seperate the plot into different facets according to name

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid( ~ name)
```

### other interesting plot

```{r}
ggplot(weather_df, aes(x = date, y = tmax)) +
  geom_point(aes(color = name))
```

using line instead

```{r}
ggplot(weather_df, aes(x = date, y = tmax)) +
  geom_line(aes(color = name))
```

seperate them

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid( ~ name)
```

if you put color parameters into aes(), and "blue" doesn't exist, it'll just create another variable "color" and make every term equals to "blue"

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

### density plot

```{r}
ggplot(weather_df, aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4)
```

### boxplot

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) +
  geom_boxplot()
```

### ridges plot

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges(scale = 0.5)
```

## last example

make a bunch of plot

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) +
  geom_boxplot()
```

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) +
  geom_violin()
```

```{r}
weather_df %>%
  filter(prcp < 100) %>%
  ggplot(aes(y = prcp, x = name)) +
  geom_boxplot()
```

## part ii

### scatterplot with parameter

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

tick makrs and labels (not that x is continuous here)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  scale_x_continuous(breaks = c(-15, 0, 15), 
                     labels = c("-15ยบ C", "0ยบ C", "15ยบ C")
  ) +
  scale_y_continuous(
    position = "right", 
    trans = "sqrt"
  )
```

### color and theme

```{r}

```